Article Title

### 2013-12-23 plos.tex
Rev30 -- nearly completed Methods. About to review with Kreso.

In this study, we sought most statistically efficient estimators of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of partial correlations between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator based on such approximations outperformed other regularized estimators. Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.  The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.
\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations in the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.  Estimation of correlation matrices can be improved by imposing some structure on the estimate. Greatest improvements are attained when the imposed structure efficiently captures real dependencies in the data. Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the performance of correlation matrix estimators that imposed different kinds of structure. The correlation structure of the estimator that led to greatest improvement comprised a sparse network of partial correlations between pairs of neurons combined with partial correlations with several common latent factors. Both components were necessary for efficient estimation. Since it leads to better estimates, we proposed that this sparse+latent structure provides a better picture of the functional connectivity than that obtained from the usual sample correlation matrix in densely sampled neural recordings. As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.

\section*{Introduction}
\paragraph{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.

@@

Given a covariance matrix estimate $C$ (not necessarily the sample covariance $C_{\sf 0}$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
\begin{equation}
    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2}
\end{equation}
Similarly, the matrix of partial correlations $P$ is computed as the negative normalized version of the \emph{precision matrix} (inverse covariance matrix) $C^{-1}$:
\begin{equation}
    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
\end{equation}
Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. Clearly, off-diagonal zeros in the precision matrix correspond to zero partial correlations.

@@

Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.  The sample estimate $C_{\sf 0}$ is unbiased ($\E{C_0}=\Sigma$) but, on average, falls far from $\Sigma$ due to its sensitivity to sampling noise.  Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.  Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate. Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.

@@

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
        The outlined square fragment is magnified in {\bf B}.
        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
    }
    \label{fig:01}}
    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
\end{figure}

\paragraph{Model selection}
Covariance selection is but one of many conceivable low-dimensional approximations of the covariance matrix. Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix. Models that accurately reflect the dominant regularities in the activity of the circuit, its so-called `functional connectivity', tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} or to \emph{dominate} other estimators. An estimator's superior efficiency can be ascertained by cross-validation.  Because different types of neural circuits have different types of regularities in their activity, different probabilistic models and covariance matrix estimators may dominate in each case. The estimator that is shown to dominate all others for a specific type of neural circuit may provide the best descriptor of the low-dimensional representation of the correlation structure of the system of interest. Its structure may then be analyzed and related to the circuit's anatomical organization.

\paragraph{Summary of findings}
In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and `sparse partial correlations with latent units' or $C_{\sf sparse+latent}$.  First, we demonstrated that, in simulations with known true low-dimensional correlation structures, regularized estimators with matching types of low-dimensional structure generally dominated over the other estimators. We then used cross-validation to determine the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex. We found that $C_{\sf sparse+latent}$ consistently dominated the other estimators.  Estimate produced by $C_{\sf sparse+latent}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.

\section*{Results}
% Results and Discussion can be combined.

\paragraph{Covariance estimation}
We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (\figref{02}{row 1}).

@@

The target of the second regularized estimator $C_{\sf factor}$, is the factor model $F =  L L^\T + \Psi$ with $d$ factors. Therefore $L$ is the $p\times d$ matrix of \emph{factor loadings} and the diagonal matrix $\Psi$ contains the independent variances of each neuron. The estimator, 
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda F,
\end{equation}
has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$. The target estimate $F$ has the structure that arises when correlated fluctuations in population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).

@@

\paragraph{Covariance estimation in neural data}
We recorded calcium activity in dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005, Katona:2012, Cotton:2013}. We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (\figref{03}{A}) on one side of the visual field of anesthetized mice. Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E} in the visual cortex on the contralateral side to the stimulus. Using acousto-optic deflectors (AOD) to steer the laser in 3D,  we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Ecker:2013}.  This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}). After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}). Only cells that produced detectable calcium activity were included in the analysis (See Methods). The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{ 03}{F and G}). The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.  Besides the high-repetition stimulus protocol used to estimate noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cell (\figref{03}{E}). Overall, 31 imaged sites from 24 animals were included in the study.

@@

Although there these concepts overlap, \emph{functional connectivity} commonly refers to the degree of correlation, synchronization, or other forms of statistical association between neuronal populations or individual neurons inferred from spontaneous activity or variability in stimulus responses,  not directly driven by external stimuli, and reflecting internal, recurrent connectivity in neural circuits.  This line of inquiry ascribes greater significance to the brain's intrinsic activity (often dismissed as `noise') \cite{Yuste:2005}. However, relatively little progress has been made in understanding the role of functional connectivity and its relationship to the anatomical connectivity.

@@

\paragraph{Cross-validation}
We used conventional 10-fold cross validation to measure cross-validation loss (Eq.~\ref{eq:cv-loss}). Briefly, each recording was split into 30 blocks of equal duration.  The 30 blocks were then grouped randomly into 10 datasets with 3 blocks in each.  This procedure ensured sufficient independence between the 10 datasets while still ensuring that each dataset included data from different parts of the recording.   Then, each dataset was used as the testing dataset with the rest of the data used for estimating the covariance matrix.

Each regularized estimator had one or two hyperparameters.   We estimated the optimal values of the hyperparameters independently in each of the 10 folds by \emph{nested cross-validation}: The 10-fold cross-validation procedure was repeated internally within each training dataset for given hyperparameter values.  The search for optimal hyperparameter values consistent of two phases: random search to find a good starting point and pattern search to find the global minimum.

\paragraph{Covariance Estimation}
Each regularized estimator starts with the sample covariance matrix $C_{\sf 0}$ computed from the training dataset and calculats their low-dimensional approximation called the target estimate.  In the last two estimators ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$), the target estimate is sufficiently flexible and is used directly.  The first two estimators ($C_{\sf diag}$ and $C_{\sf factor}$) are computed as linear mixture (shrinkage) of $C_{\sf 0}$ with the target estimate.

The diagonal matrix $D_{\sf 0}=C_{\sf 0}\circ I$ contains the sample variances.  The mean sample variance is computed as the normalized matrix trace of the sample covariance matrix:  $v_{\sf 0} = \frac 1 p \Tr(C_{\sf 0})$.

In estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag}), the diagonal matrix of regularized variances $D$ is estimated by  linearly shrinking of $D_{\sf 0}$ toward the mean variance with shrinkage intensity $\alpha$:
\begin{equation}
D = (1-\alpha)D_0 + \alpha v_{\sf 0}I 
\end{equation} 
Then the overall estimate is computed by linearly shrinking the sample covariance matrix toward $V$ with shrinkage intensity $\lambda$ (Eq.~\ref{eq:c-diag}). The {\tt corpcor} package in the programming language R also implements this estimator \cite{Schafer:2010}, although its analytical optimization of the shrinkage intensities is based on the mean square error whereas we optimized them with respect to the loss function in Eq.~\ref{eq:loss}.

Estimator $C_{\sf factor}$ is the linear shrinkage the factor analysis approximation of $C_{\sf 0}$ (Eq.~\ref{eq:c-factor}). The target $F$ by minimizing the loss $\loss{F,C_{\sf}}$, which is equivalent to maximizing the Gaussian log-likelihood with respect to $C_{\sf 0}$. To solve this minimization problem, we used the expectation-maximization (EM) algorithm.

Estimator $C_{\sf sparse}$ (Eq.~\ref{eq:c-sparse}) finds the sparse precision matrix $S$ that minimizes the loss $\loss{S^{-1},C_{\sf 0}}$.  It is computed in two steps.  First, the zero structure of the precision matrix is obtained by minimizing the $L_1$-loss:
\begin{equation}
S = \argmin\limits_{S^\prime \succ 0} \loss{{S^\prime }^{-1},C_{\sf 0}} + \lambda \|S^\prime\|_1
\end{equation}
where $S \succ 0$ is the constraint that $S$ be a positive definite matrix and $\|S\|_1$ is the element-wise $L_1$ norm of the matrix $S$. This formulation is known as \emph{graphical lasso} \cite{Friedman:2008}. To solve this minimization problem, we modified the alternative-direction method of multipliers (ADMM) algorithm developed by \cite{Ma:2013}.  Then, the $L_1$ penalty was lifted and only the non-zero elements of $S$ where fitted to minimize the loss function $\loss{S^{-1},C_{\sf 0}}$.

Estimator $C_{\sf sparse+lowrank}$ (Eq.~\ref{eq:c-sl}) finds a precision matrix $S^*$ of dimensions $(p+d)\times(p+d)$, where $d$ is the number of latent units
\begin{equation}
S^*=
\begin{pmatrix}
S_{11} & S_{12} \\
S_{12}^\T & S_{22}
\end{pmatrix}
\end{equation}
Then the covariance matrix of the observed population is 
\begin{equation}
C = \left(S_{11}-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
\end{equation}
We require that $S_{11}$ be sparse. Rather than estimate $S_{12}$ and $S_{22}$ individually, we estimate the $p\times d$ matrix $L$ such that $S_{12}S_{22}^{-1}S_{12}=LL^\T$.

\paragraph{Simulation}
TODO

\section*{Acknowledgments}
Genevera Allen

>>>>

In this study, we sought most statistically efficient estimators of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of partial correlations between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator based on such approximations outperformed other regularized estimators. Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.  The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.
\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations in the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.  Estimation of correlation matrices can be improved by imposing some structure on the estimate. Greatest improvements are attained when the imposed structure efficiently captures real dependencies in the data. Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the performance of correlation matrix estimators that imposed different kinds of structure. The correlation structure of the estimator that led to greatest improvement comprised a sparse network of partial correlations between pairs of neurons combined with partial correlations with several common latent factors. Both components were necessary for efficient estimation. Since it leads to better estimates, we proposed that this sparse+latent structure provides a better picture of the functional connectivity than that obtained from the usual sample correlation matrix in densely sampled neural recordings. As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.
\section*{Introduction}
\paragraph{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.

@@

Given a covariance matrix estimate $C$ (not necessarily the sample covariance $C_{\sf 0}$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
\begin{equation}\label{eq:precision}
    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2}
\end{equation}
Similarly, the matrix of partial correlations $P$ is computed as the negative normalized version of the \emph{precision matrix} (inverse covariance matrix) $C^{-1}$:
\begin{equation}\label{eq:partial}
    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
\end{equation}
Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. Clearly, off-diagonal zeros in the precision matrix correspond to zero partial correlations.

@@

Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.  The sample estimate $C_{\sf 0}$ is unbiased ($\E{C_0}=\Sigma$) but, on average, falls far from $\Sigma$ due to its sensitivity to sampling noise.  Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.  Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate  Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.

@@

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
        The outlined square fragment is magnified in {\bf B}.
        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
    }
    \label{fig:01}}
    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
\end{figure}
\paragraph{Model selection}
Covariance selection is but one of many conceivable low-dimensional approximations of the covariance matrix. Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix. Models that accurately reflect the dominant regularities in the activity of the circuit, its so-called `functional connectivity', tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} or to \emph{dominate} other estimators. An estimator's superior efficiency can be ascertained by cross-validation.  Because different types of neural circuits have different types of regularities in their activity, different probabilistic models and covariance matrix estimators may dominate in each case. The estimator that is shown to dominate all others for a specific type of neural circuit may provide the best descriptor of the low-dimensional representation of the correlation structure of the system of interest. Its structure may then be analyzed and related to the circuit's anatomical organization.  
\paragraph{Summary of findings}
In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and `sparse partial correlations with latent units' or $C_{\sf sparse+latent}$.  First, we demonstrated that, in simulations with known true low-dimensional correlation structures, regularized estimators with matching types of low-dimensional structure generally dominated over the other estimators. We then used cross-validation to determine the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex. We found that $C_{\sf sparse+latent}$ consistently dominated the other estimators.  Estimate produced by $C_{\sf sparse+latent}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (\figref{02}{row 1}).

@@

The target of the second regularized estimator $C_{\sf factor}$, is the factor model $LL^\T + \Psi$, where $L$ is the $p\times d$ matrix of \emph{factor loadings}, where $d$ is the number of factors, and $\Psi$ is the diagonal matrix of independent variances. The estimator, 
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (LL^\T + \Psi),
\end{equation}
has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$. The target estimate $LL^\T + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).

@@

\paragraph{Covariance estimation in neural data}
We recorded calcium activity in dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005, Katona:2012, Cotton:2013}. We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (\figref{03}{A}) on one side of the visual field of anesthetized mice. Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E} in the visual cortex on the contralateral side to the stimulus. Using acousto-optic deflectors (AOD) to steer the laser in 3D,  we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Cotton:2013}.  This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}). After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}). Only cells that produced detectable calcium activity were included in the analysis (See Methods). The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{ 03}{F and G}). The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.  Besides the high-repetition stimulus protocol used to estimate noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cell (\figref{03}{E}). Overall, 31 imaged sites from 24 animals were included in the study.

@@

In contrast to functional architecture, \emph{functional connectivity} commonly refers to the degree of correlation, synchronization, or other forms of statistical association between neuronal populations or individual neurons inferred from spontaneous activity or trial-to-trial variations,  not directly driven by external stimuli.  Functional connectivity is thought to reflect internal, recurrent connectivity in neural circuits.  This line of inquiry ascribes greater significance to the brain's intrinsic activity (often dismissed as `noise') \cite{Yuste:2005}. However, relatively little progress has been made in understanding the role of functional connectivity and its relationship to the anatomical connectivity.

@@

\paragraph{Cross-validation}
To compare the performance of the estimator against each other, we used conventional 10-fold cross validation to measure cross-validation loss (Eq.~\ref{eq:cv-loss}). Briefly, each recording was split into 30 blocks of equal duration.  The 30 blocks were then grouped randomly into 10 datasets with 3 blocks in each.  This procedure ensured sufficient independence between the 10 datasets while still ensuring that each dataset included data from different parts of the recording.   Then, each dataset was used as the testing dataset with the rest of the data used for estimating the covariance matrix.

Since each of the regularized estimators had one or two hyperparameters, we used \emph{nested cross-validation}.  The outer loop evaluated the performance of the estimators with optimal values of the hyperparameters.  The optimization of the hyperparameters was performed within the inner loop in two phases: random search to find a good starting point and pattern search to find the global minimum.  The inner cross-validation loop subdivided the training dataset from the outer loop to perform 10-fold cross-validation in order to evaluate each choice of the hyperparameter values.  Thus the size of the training dataset within the inner loop comprised 81% of the entire recording.

When cross-validation loss was not required, only the inner loop of cross-validation was used, applied to the entire dataset.  This approach was used to compute the covariance matrix estimates and their excess-loss in the simulation study (\figref{02}{\;rows 3 and 4}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:05}--\ref{fig:07}).
\paragraph{Covariance Estimation}
Within the inner loop of cross-validation, covariance matrix estimation was performed with fixed hyperparameter values provided by the search algorithm.  The computation of each regularized estimator only required the sample covariance matrix $C_{\sf 0}$ of the training dataset.

Estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag})  used two hyperparameters: the covariance shrinkage intensity $\lambda \in [0,1]$ and variance shrinkage intensity $\alpha \in [0,1]$.  The variances (the diagonal of $C_{\sf 0}$) were shrunk toward (linearly mixed with) their mean value:
\begin{equation}
D = (1-\alpha)C_{\sf 0}\circ I + \alpha \frac 1 p \Tr(C_{\sf 0}) I
\end{equation}
Then the diagonal matrix $D$ was used as the target of covariance shrinkage (Eq.~\ref{eq:c-diag}) to produce the final regularized estimate.

The {\tt corpcor} package in the programming language R also implements this estimator \cite{Schafer:2010}, although its analytical optimization of the shrinkage intensities is based on the mean squared error whereas we optimized them with respect to the loss function in Eq.~\ref{eq:loss}.

Estimator $C_{\sf factor}$ used two hyperparameters: the number of latent factors $d$ and the shrinkage intensity $\lambda \in [0, 1]$.  
The $p\times d$ factor loading matrix $L$ and individual variances $\Psi$ were computed by solving the minimization problem
\begin{equation}
(L,\Psi) = \argmin\limits_{\tilde L,\tilde\Psi} \loss{\tilde L \tilde L^\T + \tilde\Psi,C_{\sf 0}},
\end{equation}
which we solved by an expectation-maximization (EM) algorithm.  Under our chosen loss function (Eq.~\ref{eq:loss}), this is equivalent to maximum likelihood estimation of $L$ and $\Psi$ under the multivariate Gaussian distribution.  The final regularized estimate is obtained by linear shrinkage toward the factor model (Eq.~\ref{eq:c-factor}).

Estimator $C_{\sf sparse}$  has one hyperparameter $\lambda$, which regulates the sparsity of its precision matrix $S$. The precision matrix is computed in two steps: First, the zero structure $Z$ is determined by minimizing the $L_1$-penalized loss.
\begin{equation}
Z = \argmin\limits_{\tilde Z \succ 0} \loss{{\tilde Z}^{-1},C_{\sf 0}} + \lambda \|\tilde Z \|_1
\end{equation}
where $\tilde Z\succ 0$ denotes the constraint that $\tilde Z$ be a positive definite matrix and $\|\tilde Z\|_1$ is the element-wise $L_1$ norm of the matrix $\tilde Z$. This problem formulation is known as \emph{graphical lasso} \cite{Friedman:2008}. To solve this minimization problem, we modified the alternative-direction method of multipliers (ADMM) algorithm developed by \cite{Ma:2013}. 
Then, after the zero structure was determined, the remaining coefficients were fitted without penalty:
\begin{equation}
S = \argmin\limits_{\tilde S \in Z^\sharp} \loss{\tilde S,C_{\sf 0}},
\end{equation}
where $Z^\sharp$ denotes the set of positive-definite matrices with zeros in all entries where entries of $Z$ equal zero.  This step was also solved by ADMM.  Then the final estimate is the inverse of $S$ (Eq.~\ref{eq:c-sparse}). Unlike $C_{\sf diag}$ and $C_{\sf factor}$, this estimator does not include linear shrinkage: the selection of the sparsity level provides sufficient flexibility to fine tune the regularization level.

Estimator $C_{\sf sparse+lowrank}$ has two hyperparameters: the number of latent units $d$ and the sparsity level $\lambda$. It estimates the larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.  
\begin{equation}
S^\ast=
\begin{pmatrix}
S & S_{12} \\
S_{12}^\T & S_{22}
\end{pmatrix},
\end{equation}
where the $p\times p$ partition $S$ corresponds to the visible units and expresses their partial correlation structure, and $S_{12}$ and $S_{22}$ are of size $p\times d$ and $d\times d$, respectively.
Then the covariance matrix of the observed population is 
\begin{equation}
C_{\sf sparse+lowrank} = (S^\ast)^{-1} = \left(S-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
\end{equation}
The matrix $ S_{12}S_{22}^{-1}S_{12}^\T$ has rank $d$. Rather than searching for the optimal sparse structure of $S_{12}$ and $S_{22}$, an ill-posed problem, we estimated these components together as a single matrix $LL^\T$ of rank $d$, where $L$ is of size $p\times d$.
The estimate is found in two steps. First, we use the ADMM algorithm to find the zero structure of $S$ by minimizing the $L_1$-penalized loss \cite{Chandrasekaran:2010,Ma:2013}:
\begin{equation}
(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L\tilde L^\T} + \|\tilde Z\|_1
\end{equation}
Then we find the sparse and low-rank components of the precision matrix by minimizing the loss 
\begin{equation}
(S,L) = \argmin\limits_{\tilde S \in Z^\sharp,\tilde L} \loss{\tilde S-\tilde L \tilde L^\T}
\end{equation}

The MATLAB code for these computations is made publically available at {\tt http://github.com/atlab/cov-est}.

For plots in \figref{fig:05}{B and E}, \figref{fig:06}{A}, and \figref{fig:07}{A--C}, the regularized partial correlations were obtained by normalizing the precision matrix (Eq.~\ref{eq:partial}). The sparse and low-rank components in \figref{fig;05}{C} are the additive components of these normalized partial correlations.

The sparse partial correlations 
\paragraph{Simulation}

\section*{Acknowledgments}
We Genevera Allen
